---
title: "POLS 510 Final Project"
author: "Trung-Anh Nguyen"
date: "12/2/2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
library(tidyverse)
library(dplyr)
library(MASS)
library(simcf)
library(WhatIf)
library(RColorBrewer)   # For nice colors
library(pscl)
library(car) # for influence statistics and outlier detection 
library(ggstance)
library(ggthemes)
library(magrittr)
library(stargazer)
library(scales)
source("http://faculty.washington.edu/cadolph/mle/avp.R")
set.seed(1234567)
options(scipen = 999)
```


```{r setup2, include=FALSE}
## Get nice colors
brewer <- brewer.pal(9, "Set1")
purple <- brewer[4]  # Poisson on full data
red <- brewer[1]     # Poisson on non-zeros
blue <- brewer[2]    # Negative Binomial on non-zeros
orange <- brewer[5]  # Zero-inflated Negative Binomial
green <- brewer[3]   # Zero-inflated Poisson
brown <- brewer[7]    # Quasipoisson
nicegray <- "gray45"
```

```{r}
## Load data
df_1130 <- read.csv("df_1130.csv")
df_1130 <- df_1130 %>% mutate(gdp_pc2 = gdp_pc*1000,
                              pop_tot2 = pop_tot*1000000,
                              gdp = gdp_pc2*pop_tot2,
                              fdi_gdp = fdi/gdp)

#test <- df_1130 %>% dplyr::select(country, cases_cum, gov_effect, vdem_libdem, pop_tot2, pop_density, gdp_pc2, air_travel)
```


```{r histogram}
ggplot(df_1130, aes(x = cases_cum)) +
  geom_histogram() +
  scale_x_continuous(label = comma) +
  labs(x = "Cumulative case counts") +
  geom_rangeframe() +
  theme_tufte()

width <- 5
ggsave("counts_histogram.png",
       device = "png",
       width=width, 
       height=width/1.618)

ggplot(df_1130, aes(x = log(cases_cum))) +
  geom_histogram() +
  scale_x_continuous(label = comma) +
  labs(x = "Logged cumulative case counts") +
  geom_rangeframe() +
  theme_tufte()

width <- 5 
ggsave("logged_counts_histogram.png",
       device = "png",
       width=width, 
       height=width/1.618)
```

# Model 1

```{r}
# Fit model using negative binomial regression (with controls)
m1 <- cases_cum ~ gov_effect + vdem_libdem + gov_effect*vdem_libdem + log(pop_tot2) + log(pop_density) + log(gdp_pc2) + air_travel
m1_df <- extractdata(m1, df_1130, na.rm = TRUE)

nb.result1 <- glm.nb(formula=m1, data = m1_df)

pe.nb1 <- nb.result1$coefficients
vc.nb1 <- vcov(nb.result1)
se.nb1 <- sqrt(diag(vc.nb1)) 
ll.nb1 <- as.numeric(logLik(nb.result1))
theta.nb1 <- nb.result1$theta
alpha.nb1 <- 1/theta.nb1
setheta.nb1 <- nb.result1$SE.theta
print(alpha.nb1)

summary(nb.result1)
```


```{r}
## Set up counterfactuals (range over gov_effect, all else equal)
gov_effectseq <- seq(-1.9, 2.3, 0.0425)
xhyp <- cfMake(m1, m1_df, nscen = length(gov_effectseq))
for (i in 1:length(gov_effectseq)) {
  xhyp <- cfChange(xhyp, "gov_effect", x=gov_effectseq[i], scen=i)
}

# Simulate results 
sims <- 1000
simbetas.nb1 <- mvrnorm(sims, pe.nb1, vc.nb1) # draw parameters
yhyp.nb1 <- loglinsimev(xhyp, simbetas.nb1)

# Plot results with ggplot
yhyp.tidy <- yhyp.nb1 %>%
  bind_rows(
    yhyp.nb1 %>% 
      bind_rows() %>% 
      mutate(goveffect=gov_effectseq,
             model = "Model 1")
  )

ggplot(yhyp.tidy) +
  aes(y = pe,
      x = goveffect, 
      ymax = upper, 
      ymin = lower)+

  # For point estimates
  geom_line(color=blue, show.legend = FALSE) +
  
  # For confidence intervals
  geom_ribbon(alpha = 0.2, linetype = 0, show.legend = FALSE) +
  scale_y_continuous(label = comma,
                     trans = "log10") +
  labs(y = "Expected Counts of Covid-19 Cases", 
       x = "Level of Government Effectiveness") +
  theme_tufte()

width <- 5 
ggsave("counts_goveffect.png",
       device = "png",
       width=width, 
       height=width/1.618)
```


```{r}
# Counterfactuals for gov_effect at different democracy levels (-1sd, mean, +1sd)
nscen <- length(gov_effectseq)
abovemeanScen <- meanScen <- belowmeanScen <- cfMake(m1, m1_df, nscen)
for (i in 1:nscen) {
  # libdem Below Mean (loop over each gov_effect)
  belowmeanScen <- cfChange(belowmeanScen, "gov_effect", x = gov_effectseq[i], scen = i)
  belowmeanScen <- cfChange(belowmeanScen, "vdem_libdem", x = mean(m1_df$vdem_libdem) - sd(m1_df$vdem_libdem), scen = i)
  # Above Mean (loop over each gov_effect)
  abovemeanScen <- cfChange(abovemeanScen, "gov_effect", x = gov_effectseq[i], scen = i)
  abovemeanScen <- cfChange(abovemeanScen, "vdem_libdem", x = mean(m1_df$vdem_libdem) + sd(m1_df$vdem_libdem), scen = i)
  # Mean (loop over each gov_effect)
  meanScen <- cfChange(meanScen, "gov_effect", x = gov_effectseq[i], scen = i)
  meanScen <- cfChange(meanScen, "vdem_libdem", x = mean(m1_df$vdem_libdem), scen = i)
}
# Simulate expected probabilities for all scenarios
belowmeanSims <- loglinsimev(belowmeanScen, simbetas.nb1, ci=0.95)
meanSims <- loglinsimev(meanScen, simbetas.nb1, ci=0.95)
abovemeanSims <- loglinsimev(abovemeanScen, simbetas.nb1, ci=0.95)

# Plot with ggplot
# Data wrangling
belowmeanSims_tb <- belowmeanSims %>%
  bind_rows() %>%
  mutate(
    xhyp = gov_effectseq,  # add "xhyp" as a covariate
    libdem = "Less Lib.Democratic (-1sd)"  # add a column to identify which scenario
  )
meanSims_tb <- meanSims %>%
  bind_rows() %>%
  mutate(xhyp = gov_effectseq, libdem = "Average")
abovemeanSims_tb <- abovemeanSims %>%
  bind_rows() %>%
  mutate(xhyp = gov_effectseq, libdem = "More Lib.Democratic (+1sd)")
allSims_tb <- bind_rows(belowmeanSims_tb, meanSims_tb, abovemeanSims_tb)

# Visualize EVs
allSims_tb %>%
  mutate(libdem = factor(libdem, levels = c("Less Lib.Democratic (-1sd)", "Average", "More Lib.Democratic (+1sd)"))) %>%
  ggplot(aes(x = xhyp, y = pe, ymax = upper, ymin = lower, colour = libdem, fill = libdem)) +
  geom_line() +
  geom_ribbon(alpha = 0.2, linetype = 0) +
  facet_wrap(~libdem) +
  scale_y_continuous(trans = "log10",
                     label = comma) +
  labs(#title = "Predictions for countries at different liberal democracy levels",
       y = "Expected counts of Covid-19 cases", x = "Government Effectiveness") +
  theme_tufte() +
  theme(legend.position = "none",
        plot.title = element_text(size = 12,
                                  hjust = 0.5))
width <- 5.75 
ggsave("predictions_democracy_levels.png",
       device = "png",
       width=width, 
       height=width/1.618)
```


```{r}
# Low vs High democratic first differences
lowdemScen <- highdemScen <- cfMake(m1, m1_df, nscen = length(gov_effectseq))
for (i in 1:length(gov_effectseq)) {
  
  # Low Dem Scen (loop over each age)
  lowdemScen <- cfChange(lowdemScen, "gov_effect", x = gov_effectseq[i], xpre= gov_effectseq[i], scen = i)
  lowdemScen <- cfChange(lowdemScen, "vdem_libdem", 
                         x = mean(m1_df$gov_effect) - sd(m1_df$gov_effect), 
                         xpre= mean(m1_df$gov_effect) + sd(m1_df$gov_effect), scen = i)

  # High Dem (loop over each age)
  highdemScen <- cfChange(highdemScen, "gov_effect", x = gov_effectseq[i], xpre= gov_effectseq[i], scen = i)
  highdemScen <- cfChange(highdemScen, "vdem_libdem", x = mean(m1_df$gov_effect) + sd(m1_df$gov_effect), scen = i)
}

# Simulate first difference of cases with democracy level
FD <- loglinsimfd(lowdemScen, simbetas.nb1, ci=0.95)

#Plotting
fd.plot <- FD %>%
  bind_rows(
    FD %>% 
      bind_rows() %>% 
      mutate(goveffect=gov_effectseq)
  )

ggplot(fd.plot) +
  aes(y = pe,
      x = goveffect, 
      ymax = upper, 
      ymin = lower) +
  geom_line(color = "forestgreen") +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.5, linetype = 0, 
              show.legend = FALSE, fill = "gray") +
  scale_y_continuous(label = comma) +
  theme_light() +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(#title = "First Differences",
       #subtitle = "A less democratic (-1sd) compared to a more democratic (+1sd) country",
       y = "Difference in Expected Counts of Cases", 
       x = "Government Effectiveness") +
  theme_tufte() +
  theme(legend.position = "none",
        plot.title = element_text(size = 10,
                                  hjust = 0.5))
width <- 5 
ggsave("first_diff.png",
       device = "png",
       width=width, 
       height=width/1.618)
```

Difference is not significant

```{r}
## Set up counterfactuals (range over vdem_libdem, all else equal)
vdem_libdemseq <- seq(0, 1, 0.05)
xhyp2 <- cfMake(m1, m1_df, nscen = length(vdem_libdemseq))
for (i in 1:length(vdem_libdemseq))
  xhyp2 <- cfChange(xhyp2, "vdem_libdem", x=vdem_libdemseq[i], scen=i)

# Simulate results 
yhyp.nb2 <- loglinsimev(xhyp2, simbetas.nb1)

# Plot results with ggplot
yhyp.tidy2 <- yhyp.nb2 %>%
  bind_rows(
    yhyp.nb2 %>% 
      bind_rows() %>% 
      mutate(vdem_libdem=vdem_libdemseq,
             model = "Model 1")
  )

ggplot(yhyp.tidy2) +
  aes(y = pe,
      x = vdem_libdem, 
      ymax = upper, 
      ymin = lower)+

  # For point estimates
  geom_line(color=red, show.legend = FALSE) +
  
  # For confidence intervals
  geom_ribbon(alpha = 0.2, linetype = 0, show.legend = FALSE) +
  scale_y_continuous(trans = "log10",
                     label = comma) +
  labs(y = "Expected Counts of Covid-19 Cases", 
       x = "Level of liberal democracy") +
  theme_tufte() +
  theme(legend.position = "none")

width <- 5 
ggsave("counts_libdem.png",
       device = "png",
       width=width, 
       height=width/1.618)
```

```{r}
quantile(m1_df$gdp_pc2)
# Counterfactuals for gov_effect at different gdp_pc levels (-1sd, mean, +1sd)

poorScen <- averageScen <- richScen <- cfMake(m1, m1_df, nscen)

for (i in 1:nscen) {
  # Poor (loop over each gov_effect)
  poorScen <- cfChange(poorScen, "gov_effect", x = gov_effectseq[i], scen = i)
  poorScen <- cfChange(poorScen, "gdp_pc2", x = mean(m1_df$gdp_pc2) - sd(m1_df$gdp_pc2), scen = i)
  # Above Mean (loop over each gov_effect)
  richScen <- cfChange(richScen, "gov_effect", x = gov_effectseq[i], scen = i)
  richScen <- cfChange(richScen, "gdp_pc2", x = mean(m1_df$gdp_pc2) + sd(m1_df$gdp_pc2), scen = i)
  # Mean (loop over each gov_effect)
  averageScen <- cfChange(averageScen, "gov_effect", x = gov_effectseq[i], scen = i)
  averageScen <- cfChange(averageScen, "gdp_pc2", x = mean(m1_df$gdp_pc2), scen = i)
}
# Simulate expected probabilities for all scenarios
poorSims <- loglinsimev(poorScen, simbetas.nb1, ci=0.95)
averageSims <- loglinsimev(averageScen, simbetas.nb1, ci=0.95)
richSims <- loglinsimev(richScen, simbetas.nb1, ci=0.95)

# Plot with ggplot
# Data wrangling
poorSims_tb <- poorSims %>%
  bind_rows() %>%
  mutate(
    xhyp = gov_effectseq,  # add "xhyp" as a covariate
    gdp_pc2 = "Poorer (-1sd)"  # add a column to identify which scenario
  )
averageSims_tb <- averageSims %>%
  bind_rows() %>%
  mutate(xhyp = gov_effectseq, gdp_pc2 = "Average")
richSims_tb <- richSims %>%
  bind_rows() %>%
  mutate(xhyp = gov_effectseq, gdp_pc2 = "Richer (+1sd)")
allSims_tb2 <- bind_rows(poorSims_tb, averageSims_tb, richSims_tb)

# Visualize EVs
allSims_tb2 %>%
  mutate(gdp_pc2 = factor(gdp_pc2, levels = c("Poorer (-1sd)", "Average", "Richer (+1sd)"))) %>%
  ggplot(aes(x = xhyp, y = pe, ymax = upper, ymin = lower, colour = gdp_pc2, fill = gdp_pc2)) +
  geom_line() +
  geom_ribbon(alpha = 0.2, linetype = 0) +
  facet_wrap(~gdp_pc2) +
  scale_y_continuous(trans = "log10",
                     label = comma) +
  labs(title = "Predictions for countries at different GDP per capita levels",
       y = "Expected counts of Covid-19 cases (log scale)", x = "Government Effectiveness") +
  theme_tufte() +
  theme(legend.position = "none",
        plot.title = element_text(size = 12,
                                  hjust = 0.5))

width <- 5 
ggsave("prediction_gdp.png",
       device = "png",
       width=width, 
       height=width/1.618)

#Rich vs poor
poorvrich <- cfMake(m1, m1_df, nscen)
for (i in 1:nscen) {
  poorvrich <- cfChange(poorvrich, "gov_effect", x = gov_effectseq[i], scen = i)
  poorScen <- cfChange(poorvrich, "gdp_pc2", x = mean(m1_df$gdp_pc2) - sd(m1_df$gdp_pc2), scen = i)
}
# Simulate first difference of cases with democracy level
FD2 <- loglinsimfd(poorvrich, simbetas.nb1, ci=0.95)

#Plotting
fd.plot2 <- FD2 %>%
  bind_rows(
    FD2 %>% 
      bind_rows() %>% 
      mutate(goveffect=gov_effectseq)
  )

ggplot(fd.plot2) +
  aes(y = pe,
      x = goveffect, 
      ymax = upper, 
      ymin = lower) +
  geom_line(color = red) +
  geom_ribbon(aes(ymin = lower, ymax = upper),
              alpha = 0.5, linetype = 0, 
              show.legend = FALSE,
              fill = "gray") +
  scale_y_continuous(label = comma) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  labs(title = "Poorer (-1sd) compared to richer (+1sd) countries",
       y = "Difference in Expected Counts of Covid-19 Cases", 
       x = "Government Effectiveness") +
  theme_tufte() +
  theme(plot.title = element_text(size = 10,
                                  hjust = 0.5))

width <- 5 
ggsave("first_diff_2.png",
       device = "png",
       width=width, 
       height=width/1.618)
```

# Model 2

```{r}
# Model with infection variable
# infection is a dummy var for recent exp w/ SARS, MERS, Ebola (1 means 100+ cases)
#df_1130$infection <- as.factor(df_1130$infection)

m2 <- cases_cum ~ gov_effect + vdem_libdem + gov_effect*vdem_libdem + infection + gov_effect*infection + log(pop_tot2) + log(pop_density) + log(gdp_pc2) + air_travel
m2_df <- extractdata(m2, df_1130, na.rm = TRUE)

nb.result2 <- glm.nb(formula=m2, data = m2_df)

pe.nb2 <- nb.result2$coefficients
vc.nb2 <- vcov(nb.result2)
se.nb2 <- sqrt(diag(vc.nb2)) 
ll.nb2 <- as.numeric(logLik(nb.result2))
theta.nb2 <- nb.result2$theta
setheta.nb2 <- nb.result2$SE.theta
print(theta.nb2)

summary(nb.result2)

# Compare models
stargazer(nb.result1, nb.result2,
          dep.var.labels = "Cumulative COVID-19 cases",
          covariate.labels = c("Government effectiveness",
                               "Liberal democracy",
                               "Dummy for Ebola/SARS/MERS exposure(1)",
                               "ln(Total population)",
                               "ln(Population density)",
                               "ln(GDP per capita)",
                               "ln(Air travel (passengers carried))",
                               "Gov.effectiveness*Liberal democracy",
                               "Gov.effectiveness*Ebola/SARS/MERS exposure"),
          type="latex", omit="all")
```

```{r}
## Set up counterfactuals (range over gov_effect, infection)
infection0.scen <- infection1.scen <- cfMake(m2, m2_df, nscen = length(gov_effectseq))

for (i in 1:length(gov_effectseq)) {
  # For infection = 0
  infection0.scen <- cfChange(infection0.scen, "gov_effect", x = gov_effectseq[i], xpre = gov_effectseq[i], scen = i)
  infection0.scen <- cfChange(infection0.scen, "infection", x = 0, scen = i)
  # For infection = 1
  infection1.scen <- cfChange(infection1.scen, "gov_effect", x = gov_effectseq[i], xpre = gov_effectseq[i], scen = i)
  infection1.scen <- cfChange(infection1.scen, "infection", x = 3, scen = i)
}
# Simulate expected probabilities for all scenarios
simbetas.nb2 <- mvrnorm(sims, pe.nb2, vc.nb2) # draw parameters

infection0.scen <- loglinsimev(infection0.scen, simbetas.nb2, ci=0.95)
infection1.scen <- loglinsimev(infection1.scen, simbetas.nb2, ci=0.95)

# Plot with ggplot
# Data wrangling
infection0 <- infection0.scen %>%
  bind_rows() %>%
  mutate(
    xhyp = gov_effectseq,  # add "xhyp" as a covariate
    infection = "No"  # add a column to identify which scenario
  )
infection1 <- infection1.scen %>%
  bind_rows() %>%
  mutate(xhyp = gov_effectseq, infection = "Yes")

infection_tb <- bind_rows(infection0, infection1)

# Visualize EVs
infection_tb %>%
  mutate(infection= factor(infection, levels = c("No", "Yes"))) %>%
  ggplot(aes(x = xhyp, y = pe, ymax = upper, ymin = lower, color = infection, fill = infection)) +
  geom_line() +
  geom_ribbon(alpha = 0.2, linetype = 0) +
  facet_wrap(~infection) +
  scale_y_continuous(trans = "log10",
                     label = comma) +
  labs(title = "Predictions for countries with Ebola/SARS/MERS exposure",
       subtitle = "(>100 cases)",
       y = "Expected counts of Covid-19 cases", x = "Government Effectiveness") +
  theme_tufte() +
  theme(legend.position = "none",
        plot.title = element_text(size = 12))

width <- 5
ggsave("prior_exposure_M2.png",
       device = "png",
       width=width, 
       height=width/1.618)
```


# Goodness of Fit tests

## Likelihood ratio test 
```{r}
#Check number of parameters in each model
k.m1 <- length(pe.nb1) # Model 1
k.m2 <- length(pe.nb2) # Model 2
#Likelihood ratio test
(lr.test <- 2*(ll.nb2 - ll.nb1))
(lr.test.p <- pchisq(lr.test,df=(k.m2 - k.m1),lower.tail=FALSE)) #p-value for LR test
```

Difference not significant

## AIC and BIC

```{r}
#AIC
aic.m1 <- 2*k.m1 - 2*ll.nb1
aic.m2 <- 2*k.m2 - 2*ll.nb2

aic.test <- aic.m2 - aic.m1
aic.test
```

If a model is more than 2 AIC units lower than another, then it is considered significantly better than that model -> not enough evidence

```{r}
x1 <- cbind(m1_df$gov_effect, m1_df$vdem_libdem, m1_df$pop_tot2, m1_df$pop_density, m1_df$gdp_pc2, m1_df$air_travel, m1_df$gov_effect*m1_df$vdem_libdem)
x2 <- cbind(m2_df$gov_effect, m2_df$vdem_libdem, m2_df$pop_tot2, m2_df$pop_density, m2_df$gdp_pc2, m2_df$air_travel, m2_df$infection, m2_df$gov_effect*m2_df$vdem_libdem, m2_df$gov_effect*m2_df$infection)
#BIC
bic.m1 <- log(nrow(x1))*k.m1 - 2*ll.nb1
bic.m2 <- log(nrow(x2))*k.m2 - 2*ll.nb2

bic.test <- bic.m2 - bic.m1
bic.test
```

The difference between the BIC for model 2 and the BIC for model 1 is `r round(bic.test, 2)`, and according to Raftery's suggested guidelines, it demostrates strong evidence for favoring model 2 over model 1.



